*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-07 Code/Manuscript_O2.Rmd
fixed citation west et al

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focused on proteins mediating photosynthesis and protein metabolism from depths of 20 to 200 m below the ocean surface. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236). Oxygen concentration levels at the location of sampling were recorded. The methodology for sample collection and peptide analysis are described by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015].

@@

West *et al*. [@westNichePartitioningProchlorococcusPopulationsStratified1999] and Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. West *et al*.  [@westNichePartitioningProchlorococcusPopulationsStratified1999] found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)E,H), comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^ (Figure \@ref(fig:PhotoperiodPurFitsPlots)A,B).  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)I).

>>>>

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focused on proteins mediating photosynthesis and protein metabolism from depths of 20 to 200 m below the ocean surface. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise [@rollingdecktorepositoryCruiseKM1128RV2015]; original datasets in the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018; @saitoDevelopmentOceanProtein2021]. Oxygen concentration levels at the location of sampling were recorded. The methodology for sample collection and peptide analysis are described by Saito *et al*. [@saitoMultipleNutrientStresses2014; @saitoNeedlesBlueSea2015].

@@

West *et al*. [@westNichePartitioningProchlorococcusPopulations1999] and Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. West *et al*.  [@westNichePartitioningProchlorococcusPopulations1999] found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)E,H), comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^ (Figure \@ref(fig:PhotoperiodPurFitsPlots)A,B).  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR (Figure \@ref(fig:PhotoperiodPurFitsPlots)I).
